---
title: 'Course 7: Regression Models. Quiz 2'
author: "Jessica Bohning"
date: "February 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1
Consider the following data with x as the predictor and y as as the outcome.
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)
Give a P-value for the two sided hypothesis test of whether ß1 from a linear regression model is 0 or not.
```{r}
x <- c(0.61, 0.93, 0.83, 0.35, 0.54, 0.16, 0.91, 0.62, 0.62)
y <- c(0.67, 0.84, 0.6, 0.18, 0.85, 0.47, 1.1, 0.65, 0.36)

fit<-lm(y~x)
summary(fit)
```
Answer: 0.05296


## Question 2
Consider the previous problem, give the estimate of the residual standard deviation.
```{r}
summary(fit)
```
Answer: 0.223


##Question 3
In the mtcars data set, fit a linear regression model of weight (predictor) on
mpg (outcome). Get a 95% confidence interval for the expected mpg at the average
weight. What is the lower endpoint?
```{r}
data(mtcars)
x<-mtcars$wt
y<-mtcars$mpg
fit<-lm(y~x)
#Option 1
predict(fit,newdata=data.frame(x=mean(x)),interval=("confidence"))
#OPtion 2
yhat <- fit$coef[1] + fit$coef[2]*mean(x)
yhat+c(-1, 1)*qt(.975,df=fit$df)*summary(fit)$sigma/sqrt(length(y))
```
Answer: 18.991 to 21.190


## Question 4
Refer to the previous question. Read the help file for mtcars. What is the
weight coefficient interpreted as?

Answer: increasing in mpg per 1000 lbs

##Question 5
Consider again the mtcars data set and a linear regression model with mpg as
predicted by weight (1,000 lbs). A new car is coming weighing 3000 pounds. 
Construct a 95% prediction interval for its mpg. What is the upper endpoint?
```{r}
yhat<-fit$coef[1]+fit$coef[2]*3
yhat+c(-1, 1)*qt(.975,df=fit$df)*summary(fit)$sigma*sqrt(1+(1/length(y))+((3-mean(x))^2/sum((x-mean(x))^2)))
```
Answer: 14.9299 to 27.5734


##Question 6
Consider again the mtcars data set and a linear regression model with mpg as 
predicted by weight (in 1,000 lbs). A “short” ton is defined as 2,000 lbs. 
Construct a 95% confidence interval for the expected change in mpg per 1 short
ton increase in weight. Give the lower endpoint.
```{r}
fit2<-lm(y~I(x/2))
coef2<-coef(summary(fit2))
coef2[2,1]+c(-1,1)*qt(.975,df=fit$df)*coef2[2,2]
```
Answer: -12.973 to -8.405


##Question 7
If my X from a linear regression is measured in centimeters and I convert it to meters what would happen to the slope coefficient?
```{r}
x_cm<-c(5,3,8,4,1)
y<-x_cm*3
x_m<-x_cm/100
lm(y~x_cm)
lm(y~x_m)
```
Answer: x100


##Question 8
I have an outcome, Y, and a predictor, X and fit a linear regression model 
Y=ß0+ß1X+?? to obtain ß^0 and ß^1. What would be the consequence to the subsequent
slope and intercept if I were to refit the model with a new regressor, X+c for
some constant, c?
```{r}
x1<-c(3,7,34,3,7,45,267,9,5,2,7,56,4)
y<-c(1,2,78,4,345,756,32,3,8,4,9,423,89)
x2<-x1+6
lm(y~x1)
lm(y~x2)
```
Slope stayed constant while intercept shrank
Answer: 128.6887-6*.1805=127.6057


##Question 9
Refer back to the mtcars data set with mpg as an outcome and weight (wt) as the
predictor. About what is the ratio of the the sum of the squared errors, 
???ni=1(Yi-Y^i)2 when comparing a model with just an intercept (denominator) to
the model with the intercept and slope (numerator)?
```{r}
x<-mtcars$wt
y<-mtcars$mpg
fit1<-lm(y~x)
sum1<-summary(fit1)
fit2<-lm(y~x-1)
sum2<-summary(fit2)
sum2$sigma/sum1$sigma
```
Answer: About 4


##Question 10
Do the residuals always have to sum to 0 in linear regression?
Answer: They do provided an intercept is included. If not they most likely won't.


fit<-lm(mpg~wt,mtcars)
sumCoef<-summary(fit)$coefficients
sumCoef[1,1]+c(-1,1)*qt(.975,df=fit$df)*sumCoef[1,2] #CI on intercept
sumCoef[2,1]+c(-1,1)*qt(.975,df=fit$df)*sumCoef[2,2] #CI on slope
sumCoef<-summary(fit)$coefficients
sumCoef[1,1]+c(-1,1)*qt(.975,df=fit$df)*sumCoef[1,2] #CI on intercept
sumCoef[2,1]+c(-1,1)*qt(.975,df=fit$df)*sumCoef[2,2] #CI on slope