---
title: "Relationship Between MPG and Transmission"
author: "Jessica Bohning"
date: "February 24, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary
askdjjfalskjdfal;ksjdf

### Exploratory Analysis

Before starting the data analysis, let's get a bried glimpse of the data using a boxplot. 

```{r}
library(UsingR)
library(ggplot2)
data(mtcars) #an "am" of 0 = automatic and an "am" of 1 = manual
ggplot(mtcars,aes(as.character(mtcars$am),mtcars$mpg))+
        geom_boxplot(lwd=0.25,fatten=0.25,outlier.size =0.25)+
        labs(x="Automatic (0) or Manual (1) Transmission")+
        labs(y="Miles per Gallon (mpg)")
```

This graph shows that the automatic tranmission (value of 0) gets lower gas mileage than a manual transmission (1).














##Intstructions
Your report must be:

1) Written as a PDF printout of a compiled (using knitr) R markdown document.
2) Brief. Roughly the equivalent of 2 pages or less for the main text. Supporting figures in an appendix can be included up to 5 total pages including the 2 for the main report. The appendix can only include figures.
3) Include a first paragraph executive summary.
4) Upload your PDF by clicking the Upload button below the text box.