---
title: "Statistical Inference: Understanding a Normal Distribution and Examining a Tooth Growth Experiment for Statistically Signifcant Findings"
author: "Jessica Bohning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r warnings=FALSE}
#Install all libraries
library(ggplot2)
library(datasets)
```

##Overview:
In a few (2-3) sentences explain what is going to be reported on.

##Simulations
Include English explanations of the simulations you ran, with the accompanying R code. Your explanations should make clear what the R code accomplishes.

##Sample Mean versus Theoretical Mean:
Include figures with titles. In the figures, highlight the means you are comparing. Include text that explains the figures and what is shown on them, and provides appropriate numbers.

##Sample Variance versus Theoretical Variance:
Include figures (output from R) with titles. Highlight the variances you are comparing. Include text that explains your understanding of the differences of the variances.

##Distribution:
Via figures and text, explain how one can tell the distribution is approximately normal.

##Part 1: Simulation Exercise
The following will simiulate an expoential distribution with 1,000 simulations
and a rate of 0.2 and then take a random sample (with replacement) of 40

###How does the sample mean compare to the theoretical mean?
```{r cache=TRUE}
#Set up variables
set.seed(248)
lambda<- 0.2
n<-40

#run similuation 1000 times
datasimulation<-NULL
for (i in 1:1000){
        datasimulation[i]<-mean(rexp(n,lambda))
}

#samplemeans<-NULL
#Sample 40 means (with replacement)
#samplemeans<-datasimulation[sample(1:1000,40,replace=TRUE)]

#Histogram the sample means
#Theoretical mean is 1/lamba

hist(datasimulation,main="Histogram of Sample",xlab=NULL,ylab="Frequency")
abline(v=mean(datasimulation),col="red")
abline(v=1/lambda,col="blue")
legend("topright",c("Simulation Mean","Theoretical Mean"),col=c("red","blue"),lwd=1)
```

As seen in the previous histogram, the sample distribution's mean lies close to
the theoretical mean. 


###How does the sample variance compare to the theoretical variance?
```{r cache=TRUE}
samplevariance<-var(datasimulation)
theoreticalvariance<-(1/lambda)^2/n
```

The variance of the sample was `r round(samplevariance,3)` and the theoretical
variance was `r round(theoreticalvariance,3)` and this shows that they have
similar variances. 

###Is the distribution approximately normal?
If the sample mean...
```{r cache=TRUE}
set.seed(564)
ggplot(data.frame(datasimulation),aes(datasimulation))+
        geom_histogram(binwidth=.1,fill="gray",col=I("black"),
                       aes(y=..density..))+
        geom_density(col="blue")+
        stat_function(fun=dnorm,args=list(mean=1/lambda,
                                          sd=sqrt(theoreticalvariance)),
                      col="red")+
        xlab(NULL)+
        ggtitle("Distribution of Simulated Data")
```

Note: that the blue line above shows the distribution of the simulated data and
the red line shows the normal distribution

From the histogram  above, it is clear that the simulated distribution is
approximately normal


##Part 2: Basic Inferential Data Analysis
The following analysis uses the ToothGrowth data in R datasets package. In this
data, "len" is the numeric tooth length, "supp" is the supplement type (Orange
Juice or ascorbic acid), and "dose" is one of three doses of vitamin C (0.5, 1,
and 2 mg/day).

###Exploratory Analysis
The following graph splits the tooth growth data into two groups: OJ drinkers 
and ascorbic acid takers to see if there was a difference between the two.

The red dots show the mean for each dosage group (by supplement) and the black
lines show the mean for the supplement group as a whole

```{r cache=TRUE}
data(ToothGrowth)
toothmeans<-data.frame(supp=c("OJ","VC"),
                       lenmean=c(mean(ToothGrowth$len[ToothGrowth$supp=="OJ"]),
                                 mean(ToothGrowth$len[ToothGrowth$supp=="VC"])))
#Explore the data
ggplot(ToothGrowth,aes(x=dose,y=len))+geom_point()+
        facet_grid(.~supp)+
        stat_summary(fun.y="mean",geom="point",col="red",size=3)+
        geom_hline(data = toothmeans,aes(yintercept=lenmean))+
        ggtitle("Tooth Growth by Dose Method and Size")

```

The graph above shows a few things:

1. The group that received dosage in the form of OJ, had more tooth growth
2. At a dosage level of 0.5 mg/day, OJ produced more tooth growth than VC (ascorbic acid)
3. At a dosage level of 1.0 mg/day, OJ produced more tooth growth than VC
4. At a dosage level of 2.0 mg/day, OJ and VC produced approximately the same tooth growth

The next question is: are these differences statistically significant?

##Confidence Interval Testing
In the following sections, hypothesis testing will occur for the four scenarios
listed above. 

The null hypothesis for all four will be: the means for the two
samples are the same.

The alternative hypothesis is: the OJ sample is larger than the VC sample

A 95% confidence level will be used

###Hypothesis Test 1: Did the participants who took OJ have higher tooth growth than the participants that took VC?

The null hypothesis is that the means for the two samples (OJ and VC) are the same.
The alternative hypothesis is: the OJ sample is larger than the VC sample

```{r}
#Note: mean1 applies to the OJ group and mean2 applies to the VC group
mean1<-mean(ToothGrowth$len[ToothGrowth$supp=="OJ"])
mean2<-mean(ToothGrowth$len[ToothGrowth$supp=="VC"])
n1<-length(ToothGrowth$len[ToothGrowth$supp=="OJ"])
n2<-length(ToothGrowth$len[ToothGrowth$supp=="VC"])
var1<-var(ToothGrowth$len[ToothGrowth$supp=="OJ"])
var2<-var(ToothGrowth$len[ToothGrowth$supp=="VC"])

sp<-((n-1)*var1+(n2-1)*var2)/(n1+n2-2)
mean1-mean2+c(1,-1)*qt(.975,df=(n1+n2-2))*sqrt(sp)*sqrt(1/n1+1/n2)            

t.test(ToothGrowth$len[ToothGrowth$supp == "OJ"], ToothGrowth$len[ToothGrowth$supp == 
    "VC"],alternative = "greater")

```

Because the confidence interval includes zero, we accept the null hypothesis:
we cannot conclude that OJ in general is more effective at producing tooth growth

###Hypothesis Test 2: At a dose of 0.5 mg/day did the participants who took OJ have higher tooth growth than the participants that took VC?

The null hypothesis is that the means for the two samples (0.5 mg/day of OJ and 0.5 mg/day of VC) are the same.
The alternative hypothesis is: the 0.5 mg/day OJ sample is larger than the 0.5 mg/day VC sample

```{r}
#Note: mean1 applies to the OJ group and mean2 applies to the VC group
mean1<-mean(ToothGrowth$len[ToothGrowth$supp=="OJ"&ToothGrowth$dose==0.5])
mean2<-mean(ToothGrowth$len[ToothGrowth$supp=="VC"&ToothGrowth$dose==0.5])
n1<-length(ToothGrowth$len[ToothGrowth$supp=="OJ"&ToothGrowth$dose==0.5])
n2<-length(ToothGrowth$len[ToothGrowth$supp=="VC"&ToothGrowth$dose==0.5])
var1<-var(ToothGrowth$len[ToothGrowth$supp=="OJ"&ToothGrowth$dose==0.5])
var2<-var(ToothGrowth$len[ToothGrowth$supp=="VC"&ToothGrowth$dose==0.5])

sp<-((n-1)*var1+(n2-1)*var2)/(n1+n2-2)
mean1-mean2+c(1,-1)*qt(.975,df=(n1+n2-2))*sqrt(sp)*sqrt(1/n1+1/n2)            

```

Because the confidence interval includes zero, we accept the null hypothesis:
we cannot conclude that for a 0.5mg/day dose of OJ is more effective that 0.5mg/day of VC at producing tooth growth

###Hypothesis Test 3: At a dose of 1 mg/day did the participants who took OJ have higher tooth growth than the participants that took VC?

The null hypothesis is that the means for the two samples (1 mg/day of OJ and 1 mg/day of VC) are the same.
The alternative hypothesis is: the 1 mg/day OJ sample is larger than the 1 mg/day VC sample

```{r}
#Note: mean1 applies to the OJ group and mean2 applies to the VC group
mean1<-mean(ToothGrowth$len[ToothGrowth$supp=="OJ"&ToothGrowth$dose==1])
mean2<-mean(ToothGrowth$len[ToothGrowth$supp=="VC"&ToothGrowth$dose==1])
n1<-length(ToothGrowth$len[ToothGrowth$supp=="OJ"&ToothGrowth$dose==1])
n2<-length(ToothGrowth$len[ToothGrowth$supp=="VC"&ToothGrowth$dose==1])
var1<-var(ToothGrowth$len[ToothGrowth$supp=="OJ"&ToothGrowth$dose==1])
var2<-var(ToothGrowth$len[ToothGrowth$supp=="VC"&ToothGrowth$dose==1])

sp<-((n-1)*var1+(n2-1)*var2)/(n1+n2-2)
mean1-mean2+c(1,-1)*qt(.975,df=(n1+n2-2))*sqrt(sp)*sqrt(1/n1+1/n2)            

```

Because the confidence interval DO NOT include zero, we can conclude that in 95% of samples,
a 1 mg/day dose of OJ will be more effective than a 1 mg/day of VC at producing tooth growth

###Hypothesis Test 4: At a dose of 2 mg/day did the participants who took OJ have higher tooth growth than the participants that took VC?

The null hypothesis is that the means for the two samples (2 mg/day of OJ and 2 mg/day of VC) are the same.
The alternative hypothesis is: the 2 mg/day OJ sample is larger than the 2 mg/day VC sample

```{r}
#Note: mean1 applies to the OJ group and mean2 applies to the VC group
mean1<-mean(ToothGrowth$len[ToothGrowth$supp=="OJ"&ToothGrowth$dose==2])
mean2<-mean(ToothGrowth$len[ToothGrowth$supp=="VC"&ToothGrowth$dose==2])
n1<-length(ToothGrowth$len[ToothGrowth$supp=="OJ"&ToothGrowth$dose==2])
n2<-length(ToothGrowth$len[ToothGrowth$supp=="VC"&ToothGrowth$dose==2])
var1<-var(ToothGrowth$len[ToothGrowth$supp=="OJ"&ToothGrowth$dose==2])
var2<-var(ToothGrowth$len[ToothGrowth$supp=="VC"&ToothGrowth$dose==2])

sp<-((n-1)*var1+(n2-1)*var2)/(n1+n2-2)
mean1-mean2+c(1,-1)*qt(.975,df=(n1+n2-2))*sqrt(sp)*sqrt(1/n1+1/n2)            

t.test(ToothGrowth$len[ToothGrowth$supp == "OJ"], ToothGrowth$len[ToothGrowth$supp == 
    "VC"])

```

Because the confidence interval includes zero, we accept the null hypothesis:
we cannot conclude that for a 0.5mg/day dose of OJ is more effective that 0.5mg/day of VC at producing tooth growth

##Concluions
We can conclude that in 95% of samples, a 1 mg/day dose of OJ will be more effective than a 1 mg/day of VC at producing tooth growth. None of the other combinations produced statistically significant differences
